---
title: "Data_Analysis_ND_p4_nmhw"
author: Nicholas Hunt-Walker
output: html_document
---

Before we do anything, let's call the requisite libraries and go to the right working directory.
```{r}
library(ggplot2)
library(gridExtra)
library(GGally)
setwd("/Users/Nick/Documents/udacity/projects/project4/")
```


And now we load the data. Note that the column descriptions can be found [here](./Prosper Loan Data - Variable Definitions.xlsx)
```{r}
prosper <- read.csv("prosperLoanData.csv")
names(prosper)
```


Some things to know about the data set before moving forward:
- 113,937 observations, each with 81 (!!!!) variables
- Last updated 03/11/2014

---

There are five columns that contain date information. These are:
- ListingCreationDate
- ClosedDate
- DateCreditPulled
- FirstRecordedCreditLine
- LoanOriginationDate

These may be useful for later, but as they stand right now they're not usable. Let's make them datetime objects. 

Write a function that takes in a dataframe and a column of choice, and adds columns to the dataframe that splits the datetime object. I want the year, the month (maybe cyclical?), the day, as well as the date as a decimal of the year.
```{r}
split_to_columns <- function(df, input_colnum, output_colname_assoc_arr) {
  dates <- strptime(df[, input_colnum], format="%Y-%m-%d %H:%M:%S")
  df[output_colname_assoc_arr["year"]] <- as.numeric(strftime(dates, format="%Y"))
  df[output_colname_assoc_arr["month"]] <- as.numeric(strftime(dates, format="%m"))
  df[output_colname_assoc_arr["day"]] <- as.numeric(strftime(dates, format="%d"))
  df[output_colname_assoc_arr["decimal_date"]] <- df[output_colname_assoc_arr["year"]] + (as.numeric(strftime(dates, format="%j"))/366)
  
  return(df)
}
```

```{r}
#column 3
listing_creation_date_set <- c()
listing_creation_date_set["year"] <- "ListingCreationDate_year"
listing_creation_date_set["month"] <- "ListingCreationDate_month"
listing_creation_date_set["day"] <- "ListingCreationDate_day"
listing_creation_date_set["decimal_date"] <- "ListingCreationDate_decimal"

#column 7
closed_date_set <- c()
closed_date_set["year"] <- "ClosedDate_year"
closed_date_set["month"] <- "ClosedDate_month"
closed_date_set["day"] <- "ClosedDate_day"
closed_date_set["decimal_date"] <- "ClosedDate_decimal"

#column 25
credit_pulled_set <- c()
credit_pulled_set["year"] <- "DateCreditPulled_year"
credit_pulled_set["month"] <- "DateCreditPulled_month"
credit_pulled_set["day"] <- "DateCreditPulled_day"
credit_pulled_set["decimal_date"] <- "DateCreditPulled_decimal"

#column 28
first_recorded_credit_set <- c()
first_recorded_credit_set["year"] <- "FirstRecordedCreditLine_year"
first_recorded_credit_set["month"] <- "FirstRecordedCreditLine_month"
first_recorded_credit_set["day"] <- "FirstRecordedCreditLine_day"
first_recorded_credit_set["decimal_date"] <- "FirstRecordedCreditLine_decimal"

#column 65
loan_orig_date_set <- c()
loan_orig_date_set["year"] <- "LoanOriginationDate_year"
loan_orig_date_set["month"] <- "LoanOriginationDate_month"
loan_orig_date_set["day"] <- "LoanOriginationDate_day"
loan_orig_date_set["decimal_date"] <- "LoanOriginationDate_decimal"

prosper <- split_to_columns(prosper, 3, listing_creation_date_set)
prosper <- split_to_columns(prosper, 7, closed_date_set)
prosper <- split_to_columns(prosper, 25, credit_pulled_set)
prosper <- split_to_columns(prosper, 28, first_recorded_credit_set)
prosper <- split_to_columns(prosper, 65, loan_orig_date_set)
```


First Plots: Histograms of potentially interesting things
```{r}
#Listing Creation Date split by term length
ggplot(data=subset(prosper, !is.na(Term)), aes(x=ListingCreationDate_decimal)) + 
  geom_histogram(binwidth=1/12) + 
  xlab("Listing Creation Date") +
  scale_x_continuous(breaks=seq(2005, 2015)) +
  facet_wrap(~Term, ncol = 1, scales="free_y")

#Borrower APR split by term length
ggplot(data=subset(prosper, !is.na(Term)), aes(x=BorrowerAPR)) +
  geom_histogram() +
  xlab("Borrower APR (%)") +
  facet_wrap(~Term, ncol = 1, scales="free_y")

ggplot(data=subset(prosper, !is.na(Term) & !is.na(ProsperRating..Alpha.)), aes(x=ProsperRating..Alpha.)) +
  geom_histogram() +
  xlab("Credit Rating") +
  facet_wrap(~Term, ncol = 1, scales="free_y")

ggplot(data=prosper, aes(x=ListingCategory..numeric.)) +
  geom_histogram(binwidth=1) +
  xlab("Listing Category")
```

Is the stated monthly income a continuous series of numbers or is it split into categories?
```{r}
summary(prosper$StatedMonthlyIncome)
```
Looks to be continuous to me. between 0 and $1.75 million. I should be able to make a histogram of that too. This is the stated monthly income *at the time of borrowing*
```{r}
ggplot(data=prosper, aes(x=StatedMonthlyIncome)) +
  geom_histogram(binwidth=1000) +
  xlim(0, 50000) +
  xlab("Stated Monthly Income ($)") +
  facet_wrap(~Term, ncol=1, scales="free_y")
```

What about the original loan amount?
```{r}
ggplot(data=prosper, aes(x=LoanOriginalAmount)) +
  geom_histogram() +
  xlab("Loan Principal ($)")
```
Hmm.... Is there some correlation between loan amount and stated monthly income?
```{r}
with(prosper, cor.test(LoanOriginalAmount, StatedMonthlyIncome))
```
Correlation coefficient = 0.2012595. Quite lower than what I was expecting. Let's see what the scatterplot says
```{r}
ggplot(data=subset(prosper, StatedMonthlyIncome < 10000), aes(y=LoanOriginalAmount, x=StatedMonthlyIncome)) +
  geom_point(position="jitter", alpha=0.2, aes(color=factor(Term))) +
  xlab("Stated Monthly Income ($)") +
  ylab("Loan Principal ($)") 
```

I wonder if there's any visible difference when coloring on Borrower APR instead of just term.
```{r}
ggplot(data=subset(prosper, (StatedMonthlyIncome < 10000) & (EmploymentStatus != "")),
       aes(y=LoanOriginalAmount, x=StatedMonthlyIncome)) +
  geom_point(position="jitter", alpha=0.2, aes(color=factor(EmploymentStatus))) +
  xlab("Stated Monthly Income ($)") +
  ylab("Loan Principal ($)") 
```

Somewhat. Smaller loans for those employed part time. A ton of borrowers are employed or self-employed though. Let's see the employment statuses of borrowers more explicitly.
```{r}
ggplot(data=subset(prosper, (EmploymentStatus != "") & (EmploymentStatus != "Not available")), aes(x=EmploymentStatus)) +
  geom_bar() +
  xlab("Employment Status") +
  scale_y_log10()
```

This could be an interesting thing to look at. Is there a split on Term?
```{r}
ggplot(data=subset(prosper, (EmploymentStatus != "") & (EmploymentStatus != "Not available")), aes(x=EmploymentStatus)) +
  geom_bar(aes(fill=factor(Term))) +
  xlab("Employment Status") +
  scale_y_log10()
```
Oooh this is money right here. 

Here's another question then, is there a significant difference between the ways different employment statuses borrow over time?
```{r}
ggplot(data=subset(prosper, !is.na(LoanOriginalAmount) & (EmploymentStatus != "")), aes(x=ListingCreationDate_decimal, y=LoanOriginalAmount)) + 
  geom_point(aes(color=factor(EmploymentStatus)), alpha=0.1, position="jitter") + 
  xlab("Listing Creation Date") +
  scale_x_continuous(breaks=seq(2005, 2015)) +
  scale_y_sqrt()
```

The employed and full-time borrowers wash out everyone else, so let's see what it looks like without them.
```{r}
ggplot(data=subset(prosper, !is.na(LoanOriginalAmount) & (EmploymentStatus != "") & (EmploymentStatus != "Not available") & (EmploymentStatus != "Employed") & (EmploymentStatus != "Full-time")), aes(x=ListingCreationDate_decimal, y=LoanOriginalAmount)) + 
  geom_point(aes(color=factor(EmploymentStatus)), alpha=0.2, position="jitter") + 
  xlab("Listing Creation Date") +
  scale_x_continuous(breaks=seq(2005, 2015))
```


Onto other things. There's a variable for the percent of the listing funded. I suspect there may be a correlation between the PercentFunded and the amount asked for.
```{r}
with(prosper, cor.test(LoanOriginalAmount, PercentFunded))
```
-0.01024815 is basically no correlation at all. Disappointment. 

Let's see what that PercentFunded parameter looks like
```{r}
ggplot(data=subset(prosper, !is.na(PercentFunded)), aes(x=PercentFunded)) +
  geom_histogram(binwidth=0.1, origin=-0.05) +
  xlim(0.6, 1.1)

ggplot(data=subset(prosper, !is.na(PercentFunded)), aes(x=PercentFunded)) +
  geom_histogram(binwidth=0.1, origin=-0.05) +
  xlim(0.6, 1.1) + scale_y_log10()
```
Oh that's why. Because the vaaaaast majority of loans are fully funded. That was useless and enlightening all at once. 

Let's quantify just how many are fully funded
```{r}
sum(with(prosper, (PercentFunded == 1.0) & !is.na(PercentFunded))) / sum(with(prosper, !is.na(PercentFunded))) * 100
```
99.236%! Yeah, we're not going to get any meaningful correlations out of this one.











