---
title: "lesson3"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)

```

Histogram of people's birthdays
```{r}
setwd("/Users/Nick/Documents/udacity/projects/data_analysis_with_r/")
list.files()
pf = read.delim("pseudo_facebook.tsv", sep="\t")
names(pf)
qplot(x=dob_day, data=pf) + 
  scale_x_discrete(breaks=1:31)
```

Faceting
```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram() + 
  scale_x_discrete(breaks = 1:31) + 
  facet_wrap(~dob_month, ncol=4)
```

Adjusting axes
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x = friend_count)) + 
  geom_histogram() + scale_x_continuous(limits = c(0, 1000), breaks=seq(0, 1000, 10)) + 
  facet_wrap(~gender)
```

Table based on data
Table of summarized friend count statistics by gender
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

Tenure
```{r}
ggplot(data=pf, aes(x=tenure/365)) +
  geom_histogram(color=I("black"), fill=I("#099DD9")) + 
  scale_x_continuous(breaks=seq(1,7,1)) + 
  xlab("Number of years using Facebook") + ylab("Number of Users")
```

Ages
```{r}
ggplot(data=subset(pf, !is.na(gender)), aes(x=age, fill=gender)) +
  geom_histogram(color=I("white"), binwidth=2) +
  xlab("Age of User (years)") + ylab("Number of Users") +
  facet_wrap(~gender)
```

Transforming data
```{r}
ggplot(data=pf, aes(x=friend_count)) + geom_histogram()
```

Multiple plots
```{r}
plot1 = ggplot(data=pf, aes(x=friend_count)) + geom_histogram()
plot2 = ggplot(data=pf, aes(x=friend_count)) + geom_histogram() + scale_x_log10()
plot3 = ggplot(data=pf, aes(x=friend_count)) + geom_histogram() + scale_x_sqrt()
grid.arrange(plot1, plot2, plot3, ncol=1)
```

Multiple data on one plot
```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 100)) +
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```

Multiple data on one plot, this time with www_likes
```{r}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
    scale_x_log10() +
  xlab('Likes Online') + 
  ylab('Percentage of users with that friend count')

by(pf$www_likes, pf$gender, sum)
```

Box Plot
```{r}
ggplot(aes(x = gender, y=friend_count), data = subset(pf, !is.na(gender))) + 
  geom_boxplot() + scale_y_continuous(limits=c(0,250))
```

---
## Lesson 4
---

Who initiated more friendships?
```{r}
by(pf$friendships_initiated, pf$gender, summary)
```

Mobile Checkins
```{r}
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/(sum(pf$mobile_check_in == 1) + sum(pf$mobile_check_in == 0)) * 100
```

---
## Lesson 5
---

Third Qualitative Variable
```{r}
ggplot(aes(x=gender, y=age), data=subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = median, geom="point", shape=4)

ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(gender))) + 
  geom_line(aes(color=gender), stat="summary", fun.y=median)
```


Create a dataframe showing the mean and median friend count by age and gender.
```{r}
library(dplyr)
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>% 
  group_by(age, gender) %>% 
  summarise(mean_friend_count=mean(friend_count),
            median_friend_count = median(friend_count),
            n = n())
pf.fc_by_age_gender
```


Now, create a line graph showing the median friend count over the ages
for each gender. Be sure to use the data frame you just created, pf.fc_by_age_gender.
```{r}
ggplot(data=pf.fc_by_age_gender, aes(x=age, y=median_friend_count)) +
  geom_line(aes(color=gender))
```

We now want to reshape this table where we have one row for age, and one column for each gender.
Left of the ~ goes the value we want to keep as is, to the right of the ~ goes the values we want as columns. Value.var holds the key measurements of our dataframe.
```{r}
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = "median_friend_count")
```

Now plot the ratio of female-to-male median friend count. Include a horizontal
line at ratio = 1
```{r}
ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female / male)) + geom_line() + 
  geom_hline(aes(yintercept = 1), linetype = 2, alpha = 0.3)
```

Let's look at the year the user first joined facebook using the variable "tenure" and 2014 as the reference year. Add this to the pf data frame as the variable "year_joined"
```{r}
pf$year_joined = floor(2014 - pf$tenure / 365)
```

Cut a variable to make discrete variables from continuous or numerical ones. Cut year_joined to create 4 bins of users:
2004-2009
2009-2011
2011-2012
2012-2014
```{r}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```


Create a plot for friend_count vs age so each year_joined.bucket is a line
tracking the median user friend_count across age. This means you should have
four different lines on your plot. Exclude users whose year_joined.bucket is NA.
```{r}
ggplot(data=subset(pf, !is.na(year_joined)), aes(x = age, y = friend_count)) +
  geom_line(aes(color=year_joined.bucket), fun.y=median, stat="summary")
```


Change the above plot to show the means instead of the median, and also plot the
grand mean of the overall sample.
```{r}
ggplot(data=subset(pf, !is.na(year_joined)), aes(x = age, y = friend_count)) +
  geom_line(aes(color=year_joined.bucket), fun.y=mean, stat="summary") +
  geom_line(fun.y=mean, stat="summary", linetype=2)
```


Friending rate. Friend count vs Tenure. How many friends does a user have for 
each day since they've started using Facebook? Create a summary of this rate 
showing how many friends a user has for each day they've been using facebook. 
Make sure tenure > 0. What's the median rate? What's the maximum rate?
```{r}
friend_rate = with(subset(pf, tenure > 0), friend_count/tenure)
summary(friend_rate)
```


Create a line graph of mean friendships initiated per day of tenure vs 
tenure (tenure > 0), colored by year_joined.bucket. 
```{r}
ggplot(data=subset(pf, tenure > 0), 
       aes(x=7 * round(tenure / 7), y=friendships_initiated/tenure)) +
  geom_smooth(aes(color=year_joined.bucket))
  #geom_line(stat="summary", fun.y=mean, aes(color=year_joined.bucket))
  
```


Scatterplot Matrix
```{r}
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```

