---
title: "PS3_birthdays"
output: html_document
---

```{r}
library(chron)
birthdays = read.csv("facebook_birthdays.csv")
```

Reformat dates and add columns
```{r}
birthdays$Start <- strptime(birthdays$Start, format="%m/%d/%Y %H:%M")
birthdays$Day <- strftime(birthdays$Start, format="%d")
birthdays$Month <- strftime(birthdays$Start, format="%m")
birthdays$Year <- strftime(birthdays$Start, format="%Y")
birthdays$Bool_col <- TRUE
birthdays$Weekday <- strftime(birthdays$Start, format="%A")
birthdays$Yearday <- strftime(birthdays$Start, format="%j")
```

How many people share my birthday?
```{r}
sum(strftime(birthdays$Start, format="%m-%d") == "04-18")
```
answer: 3 people

Which month has the most birthdays?
```{r}
by(birthdays$Bool_col, birthdays$Month, sum)
```
answer: July

How many birthdays are in each month?
```{r}
by(birthdays$Bool_col, birthdays$Month, sum)
ggplot(data=birthdays, aes(x=birthdays$Month)) + geom_histogram() +
  scale_x_discrete(labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                            "Aug", "Sep", "Oct", "Nov", "Dec")) +
  xlab("Birth Month") + ylab("Number of People")
```
January: 32
February: 33
March: 35
April: 41
May: 41
June: 38
July: 49
August: 39
September: 36
October: 40
November: 37
December: 36

Which day of the year has the most number of birthdays?
```{r}
split_by_weekday = by(birthdays$Bool_col, birthdays$Weekday, sum)
split_by_weekday[split_by_weekday == max(split_by_weekday)]
split_by_day = by(birthdays$Bool_col, birthdays$Day, sum)
split_by_day[split_by_day == max(split_by_day)]
```
Monday: 78 birthdays
6th of the month: 25 birthdays

Do you have at least 365 friends that have birthdays on everyday
of the year?
```{r}
split_by_yearday = by(birthdays$Bool_col, birthdays$Yearday, sum)
split_by_yearday[split_by_yearday > 0]
```
At least 365 friends? yes
Birthdays on every day? no

