---
title: "PS3_gapminder"
output: html_document
---
Once you've completed your investigation, create a post in the discussions that includes:
       1. any questions you answered, your observations, and summary statistics
       2. snippets of code that created the plots
       3. links to the images of your plots
---

The Gapminder website contains over 500 data sets with information about
the world's population. Your task is to download a data set of your choice
and create 2-5 plots that make use of the techniques from Lesson 3.

You might use a simple histogram, a boxplot split over a categorical variable,
or a frequency polygon. The choice is yours!

I've chosen arms imports and exports worldwide. They're both .xlsx files, so lets 
start with a few preliminaries.

```{r}
library(ggplot2)
library(gridExtra)
setwd("/Users/Nick/Documents/udacity/projects/data_analysis_with_r/")
```

I've decided on two data sets actually. They're related, so I'm considering them to be one data set. They are:
- Arms exports (US$, inflation-adjusted)
- Arms imports (US$, inflation-adjusted)

Let's load in each data set. I'm going to want to compare the two. Note that because they're saved from .xlsx files, their columns (years) are all prefixed with "X". I'm removing that. I'm also changing the "Arms.exports..constant.1990.US.." column to "country", as these are the data in that column. Lastly, we're transposing because we want the data to be per country, not per year.
```{r}
arms_in <- read.csv("arms_imports.csv")
names(arms_in) <- gsub("X", "", names(arms_in))
names(arms_in) <- sub("^Arms.imports..constant.1990.US..","year", names(arms_in))
arms_in <- t(arms_in)

arms_out <- read.csv("arms_exports.csv")
names(arms_out) <- gsub("X", "", names(arms_out))
names(arms_out) <- sub("Arms.exports..constant.1990.US..","year", names(arms_out))
arms_out <- t(arms_out)
```


Separate out 5 first-world nations
- United States
- United Kingdom
- Switzerland
- China
- Germany

and 5 developing nations 
- Jordan
- Korea, Rep.
- Libya
- Pakistan
- Vietnam

WITH good data coverage across all years. Then combine all import and export data for these years into one data frame
```{r}
years = c(1960:2011)
usa_imp = as.numeric(arms_in[2:53, 203])
uk_imp = as.numeric(arms_in[2:53, 202])
swiss_imp = as.numeric(arms_in[2:53, 185])
china_imp = as.numeric(arms_in[2:53, 41])
ger_imp = as.numeric(arms_in[2:53, 71])
jordan_imp = as.numeric(arms_in[2:53, 97])
korea_imp = as.numeric(arms_in[2:53, 102])
libya_imp = as.numeric(arms_in[2:53, 111])
pak_imp = as.numeric(arms_in[2:53, 147])
viet_imp = as.numeric(arms_in[2:53, 208])


usa_exp = as.numeric(arms_out[2:53, 203])
uk_exp = as.numeric(arms_out[2:53, 202])
swiss_exp = as.numeric(arms_out[2:53, 185])
china_exp = as.numeric(arms_out[2:53, 41])
ger_exp = as.numeric(arms_out[2:53, 71])
jordan_exp = as.numeric(arms_out[2:53, 97])
korea_exp = as.numeric(arms_out[2:53, 102])
libya_exp = as.numeric(arms_out[2:53, 111])
pak_exp = as.numeric(arms_out[2:53, 147])
viet_exp = as.numeric(arms_out[2:53, 208])

first_world = data.frame(years, usa_imp, usa_exp, uk_imp, uk_exp, swiss_imp,
                         swiss_exp, china_imp, china_exp, ger_imp, ger_exp)
third_world = data.frame(years, jordan_imp, jordan_exp, korea_imp, korea_exp,
                         libya_imp, libya_exp, pak_imp, pak_exp, viet_imp, viet_exp)
```


First plots: line graphs of arms imports/exports for the first-world countries over time
```{r}
library(grid)
plot1 = ggplot(data=first_world, aes(x=years)) + geom_line(aes(y=usa_exp/1E9)) +
  geom_line(aes(y=uk_exp/1E9), color="red") +
  geom_line(aes(y=swiss_exp/1E9), color="green") +
  geom_line(aes(y=china_exp/1E9), color="blue") +
  geom_line(aes(y=ger_exp/1E9), color="orange") +
  scale_y_continuous(limits=c(0, 17.5)) + xlab("Year") +
  ylab("Arms Exports per $US billion") + 
  theme(legend.position="none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(1, 1, -0.5, 1), "cm"))

plot2 = ggplot(data=first_world, aes(x=years)) + geom_line(aes(y=usa_imp/1E9)) +
  geom_line(aes(y=uk_imp/1E9), color="red") +
  geom_line(aes(y=swiss_imp/1E9), color="green") +
  geom_line(aes(y=china_imp/1E9), color="blue") +
  geom_line(aes(y=ger_imp/1E9), color="orange") +
  scale_y_continuous(limits=c(0, 17.5)) +
  ylab("Arms Imports per $US billion") + 
  theme(legend.position="none", 
        plot.margin = unit(c(-0.05, 1, 1, 1), "cm"))

grid.arrange(plot1, plot2, ncol=1)
#ggsave("first_world_arms_inout.png")
```

Second set: Same plots, but for third-world countries
```{r}
plot1 = ggplot(data=third_world, aes(x=years)) + geom_line(aes(y=jordan_exp/1E6)) +
  geom_line(aes(y=korea_exp/1E6), color="red") +
  geom_line(aes(y=libya_exp/1E6), color="green") +
  geom_line(aes(y=pak_exp/1E6), color="blue") +
  geom_line(aes(y=viet_exp/1E6), color="orange") +
  scale_y_continuous(limits=c(0, 500)) + xlab("Year") +
  ylab("Arms Exports per $US million") + 
  theme(legend.position="none", axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(1, 1, -0.5, 1), "cm"))

plot2 = ggplot(data=third_world, aes(x=years)) + geom_line(aes(y=jordan_imp/1E9)) +
  geom_line(aes(y=korea_imp/1E9), color="red") +
  geom_line(aes(y=libya_imp/1E9), color="green") +
  geom_line(aes(y=pak_imp/1E9), color="blue") +
  geom_line(aes(y=viet_imp/1E9), color="orange") +
  scale_y_continuous(limits=c(0, 5)) +
  ylab("Arms Imports per $US billion") + 
  theme(legend.position="none", 
        plot.margin = unit(c(-0.05, 1, 1, 1), "cm"))

grid.arrange(plot1, plot2, ncol=1)
# ggsave("third_world_arms_inout.png")
```


Third: It looks as if the US is the biggest seller of arms since 1960. Is that true?
```{r}
ggplot(data = first_world) + 
  geom_boxplot(aes(x=years, y=usa_exp/1E9), fill="black", color="white") +
  geom_boxplot(aes(x=years + 50, y=uk_exp/1E9), fill="red", color="white") +
  geom_boxplot(aes(x=years + 100, y=swiss_exp/1E9), fill="green", color="white") +
  geom_boxplot(aes(x=years + 150, y=china_exp/1E9), fill="blue", color="white") +
  geom_boxplot(aes(x=years + 200, y=ger_exp/1E9), fill="orange", color="white") +
  ylab("Arms Exports per $US billion") +
  xlab("Nation") + 
  scale_x_continuous(breaks=c(1985, 2035, 2085, 2135, 2185), 
                     labels=c("USA", "UK", "Switzerland", "China", "Germany"))
# ggsave("first_world_arms_boxplot.png")
```

Fourth: Let's look at those imports for third world countries.
```{r}
ggplot(data = third_world) + 
  geom_boxplot(aes(x=years, y=jordan_imp/1E9), fill="black", color="white") +
  geom_boxplot(aes(x=years + 50, y=korea_imp/1E9), fill="red", color="white") +
  geom_boxplot(aes(x=years + 100, y=libya_imp/1E9), fill="green", color="white") +
  geom_boxplot(aes(x=years + 150, y=pak_imp/1E9), fill="blue", color="white") +
  geom_boxplot(aes(x=years + 200, y=viet_imp/1E9), fill="orange", color="white") +
  ylab("Arms Imports per $US billion") +
  xlab("Nation") + 
  scale_x_continuous(breaks=c(1985, 2035, 2085, 2135, 2185), 
                     labels=c("Jordan", "S. Korea", "Libya", "Pakistan", "Vietnam"))
# ggsave("third_world_arms_boxplot.png")
```

---
# Problem Set 4
---
Let's use multiple data sets more wisely. We have already arms imports and arms exports. We can also get GDP data (in USD 2000 which we can convert to 1990 or vice versa with a factor of 1.317521041), and others. We can merge them into a data frame with columns like so:
- country
- year
- arms imports USD 1990 (or 2000)
- arms exports USD 1990 (or 2000)
- GDP USD 1990 (or 2000)
- Life Expectancy (years)

Let's first load and format the appropriate data sets
```{r}
arms_in <- read.csv("arms_imports.csv")
arms_out <- read.csv("arms_exports.csv")
gdp_2000 <- read.csv("gdp_per_capita_usd2000.csv")
life_expectancy <- read.csv("life_expectancy_at_birth.csv")
unemp_fems <- read.csv("females_unemployment_pct.csv")
unemp_males <- read.csv("males_unemployment_pct.csv")

names(arms_in) <- gsub("X", "", names(arms_in))
names(arms_out) <- gsub("X", "", names(arms_out))
names(gdp_2000) <- gsub("X", "", names(gdp_2000))
names(life_expectancy) <- gsub("X", "", names(life_expectancy))
names(unemp_fems) <- gsub("X", "", names(unemp_fems))
names(unemp_males) <- gsub("X", "", names(unemp_males))

names(arms_in) <- sub("^Arms.imports..constant.1990.US..",
                      "country", names(arms_in))
names(arms_out) <- sub("^Arms.exports..constant.1990.US..",
                       "country", names(arms_out))
names(gdp_2000) <- sub("^Income.per.person..fixed.2000.US..",
                       "country", names(gdp_2000)) 
names(life_expectancy) <- sub("^Life.expectancy.with.projections..Yellow.is.IHME",
                              "country", names(life_expectancy))
names(unemp_fems) <- sub("^Female.15..unemployment....",
                              "country", names(unemp_fems))
names(unemp_males) <- sub("^Male.15..unemployment....",
                              "country", names(unemp_males))
```


Let's try to have a data frame combining information from the year 2000
for all countries
```{r}
all_info = arms_in[c("country", "2000")]
names(all_info)[2] <- "imports_2000"

exports_dum = arms_out[, c("country", "2000")]
all_info = merge(all_info, exports_dum, by="country")
names(all_info)[3] <- "exports_2000"

gdp_dum = gdp_2000[, c("country", "2000")]
all_info = merge(all_info, gdp_dum, by="country")
names(all_info)[4] <- "gdp_2000"

life_dum = life_expectancy[, c("country", "2000")]
all_info = merge(all_info, life_dum, by="country")
names(all_info)[5] <- "life_expectancy_2000"

fems_dum = unemp_fems[, c("country", "2000")]
all_info = merge(all_info, fems_dum, by="country")
names(all_info)[6] <- "female_unemployment_2000"

males_dum = unemp_males[, c("country", "2000")]
all_info = merge(all_info, males_dum, by="country")
names(all_info)[7] <- "male_unemployment_2000"
```

```{r}
ggplot(data=all_info, aes(x=gdp_2000, y=life_expectancy_2000)) + 
  geom_point() +
  xlab("GDP in 2000 (USD 2000)") + ylab("Life Expectancy (years)")

ggplot(data=all_info, aes(x=life_expectancy_2000, y=female_unemployment_2000)) + 
  geom_point() +
  ylab("Female Unemployment in 2000 (%)") + xlab("Life Expectancy (years)")
```


---
# Problem Set 5
---
Adding two new data sets onto the overall set. Let's reconstruct it
for the year 2002 instead of the year 2000.
```{r}
arms_in <- read.csv("arms_imports.csv")
arms_out <- read.csv("arms_exports.csv")
gdp_2000 <- read.csv("gdp_per_capita_usd2000.csv")
life_expectancy <- read.csv("life_expectancy_at_birth.csv")
unemp_fems <- read.csv("females_unemployment_pct.csv")
unemp_males <- read.csv("males_unemployment_pct.csv")
suicides <- read.csv("suicides_per_100k.csv")
healthcare <- read.csv("health_expenditure_pct_gdp.csv")

names(arms_in) <- gsub("X", "", names(arms_in))
names(arms_out) <- gsub("X", "", names(arms_out))
names(gdp_2000) <- gsub("X", "", names(gdp_2000))
names(life_expectancy) <- gsub("X", "", names(life_expectancy))
names(unemp_fems) <- gsub("X", "", names(unemp_fems))
names(unemp_males) <- gsub("X", "", names(unemp_males))
names(suicides) <- gsub("X", "", names(suicides))
names(healthcare) <- gsub("X", "", names(healthcare))

names(arms_in) <- sub("^Arms.imports..constant.1990.US..",
                      "country", names(arms_in))
names(arms_out) <- sub("^Arms.exports..constant.1990.US..",
                       "country", names(arms_out))
names(gdp_2000) <- sub("^Income.per.person..fixed.2000.US..",
                       "country", names(gdp_2000)) 
names(life_expectancy) <- sub("^Life.expectancy.with.projections..Yellow.is.IHME",
                              "country", names(life_expectancy))
names(unemp_fems) <- sub("^Female.15..unemployment....",
                              "country", names(unemp_fems))
names(unemp_males) <- sub("^Male.15..unemployment....",
                              "country", names(unemp_males))
names(suicides) <- sub("^Suicide..age.adjusted..per.100.000.standard.population",
                       "country", names(suicides)) 
names(healthcare) <- sub("^Total.expenditure.on.health.as.percentage.of.GDP..gross.domestic.product.",
                       "country", names(healthcare)) 
```

pick year 2002. Inflation coefficient going from USD 1990 -> 2002 is 1.37643458
```{r}
all_info = arms_in[c("country", "2002")]
names(all_info)[2] <- "imports_2002"

exports_dum = arms_out[, c("country", "2002")]
all_info = merge(all_info, exports_dum, by="country")
names(all_info)[3] <- "exports_2002"

gdp_dum = gdp_2000[, c("country", "2002")]
all_info = merge(all_info, gdp_dum, by="country")
names(all_info)[4] <- "gdp_2002"

life_dum = life_expectancy[, c("country", "2002")]
all_info = merge(all_info, life_dum, by="country")
names(all_info)[5] <- "life_expectancy_2002"

suicides_dum = suicides[, c("country", "2002")]
all_info = merge(all_info, suicides_dum, by="country")
names(all_info)[6] <- "suicides_per_100k_2002"

health_dum = healthcare[, c("country", "2002")]
all_info = merge(all_info, health_dum, by="country")
names(all_info)[7] <- "healthcare_pct_gdp_2002"
```

First, let's put suicides and healthcare into buckets
```{r}
all_info$suicide_buckets = cut(all_info$suicides_per_100k_2002, 
                               quantile(all_info$suicides_per_100k_2002,
                                        probs=c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)))
all_info$health_buckets = cut(all_info$healthcare_pct_gdp_2002,
                              c(0, 5, 10, 15, 20))
```


Lets see arms imports vs life expectancy, colored by suicides
```{r}
plt1 = ggplot(data=subset(all_info, imports_2002 > 0),
       aes(x=life_expectancy_2002, y=imports_2002)) + 
  geom_jitter(aes(color=health_buckets)) +
  ylab("Arms Imports 2002 (USD - 1990)") +
  scale_y_log10()
ggsave("arms_imports_vs_lifetime_healthcare.png")

plt2 = ggplot(data=subset(all_info, imports_2002 > 0),
       aes(x=life_expectancy_2002, y=imports_2002)) + 
  geom_jitter(aes(color=suicide_buckets)) +
  xlab("Life Expectancy (years, 2002)") +
  ylab("Arms Imports 2002 (USD - 1990)") +
  scale_y_log10()

grid.arrange(plt1, plt2, nrow=2)
ggsave("arms_imports_vs_lifetime_suicides.png")
```


Now let's see the suicide rate vs gdp, colored by percent of gdp
spent on healthcare
```{r}
ggplot(data=all_info, 
       aes(x=gdp_2002, y=life_expectancy_2002)) + 
  geom_jitter(aes(color=health_buckets)) +
  ylab("Life Expectancy (years, 2002)") +
  xlab("GDP Per Capita in 2002 (USD - 2000)") +
  scale_x_log10(breaks=c(1E2, 3.16E2, 1E3, 3.16E3, 1E4, 3.16E4, 1E5, 3.16E5)) + geom_smooth()
ggsave("lifetime_vs_gdp.png")
```

