---
title: "PS4_diamonds"
output: html_document
---

Load the diamonds data set, as well as ggplot2
```{r}
setwd("/Users/Nick/Documents/udacity/projects/data_analysis_with_r/")
library(ggplot2)
data(diamonds)
```


Create a scatterplot of price vs x for the diamonds data set
```{r}
ggplot(data=diamonds, aes(x=x, y=price)) + geom_point()
```


Testing correlations between price and x/y/z
```{r}
with(diamonds, cor.test(x, price))
with(diamonds, cor.test(y, price))
with(diamonds, cor.test(z, price))
```
- between price and x? 0.88
- between price and y? 0.87
- between price and z? 0.86


Create a simple scatter plot of price vs depth. Add to the code some transparency in the dots (1/100) and mark the x-axis every 2 units.
```{r}
ggplot(data=diamonds, aes(x=depth, y=price)) + geom_point(alpha=0.01) +
  scale_x_continuous(breaks=seq(40, 80, 2))
```
The above plot places the vast majority of diamonds between depths of 60 and 64.


What's the correlation of depth vs. price?
```{r}
with(diamonds, cor.test(depth, price))
```
The correlation between the two is -0.01. This would be a horrible predictive characteristic of diamond price.


Create a scatterplot of price vs carat, and omit the top 1% of price and carat values.
```{r}
ggplot(data=diamonds, aes(x=carat, y=price)) + geom_point() + 
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```


Create a scatterplot of price vs volume (x * y * z). In the process, create a new variable for volume in the diamonds data frame.
```{r}
diamonds$volume = diamonds$x * diamonds$y * diamonds$z
ggplot(data=diamonds, aes(x=volume, y=price)) + geom_point()
```
The x-axis needs to be rescaled to exclude the worst outliers (like 3?). Generally though, price trends upward with diamond volume.


What's the correlation of price and volume? Exclude diamonds that have a volume of 0 or that are greater than or equal to 800.
```{r}
with(subset(diamonds, (diamonds$volume > 0) & (diamonds$volume < 800)), cor.test(volume, price))
```
Correlation: 0.9235455


Subset the data to exclude diamonds with a volume greater tha or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.

Do you think this would be a useful model to estimate the price of diamonds? Why or why not?
```{r}
diam_sub = subset(diamonds, (diamonds$volume > 0) & (diamonds$volume < 800))
ggplot(data=diam_sub, aes(x=volume, y=price)) + geom_point(alpha=0.01) +
  geom_smooth()
```
This is definitely not a useful model beyond a volume of about 400. Perhaps even before then.


Use the function dplyr package to create a new data frame containing info on diamonds by clarity. Name the data frame "diamondsByClarity". The data should contain the following variables in this order:
- mean_price
- median_price
- min_price
- max_price
- n
where "n" is the number of diamonds in each level of clarity
```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n())
diamondsByClarity
```


`diamonds_mp_by_clarity` and `diamonds_mp_by_color` are summary data frames with the mean price by clarity and color. Create two bar plots on one output image using the grid.arrange() function from the package gridExtra
```{r}
library(gridExtra)
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

plot1 = ggplot(data=diamonds_mp_by_clarity, aes(factor(clarity), y=mean_price)) + geom_bar(stat="identity")
plot2 = ggplot(data=diamonds_mp_by_color, aes(x=factor(color), y=mean_price)) + geom_bar(stat="identity")
grid.arrange(plot1, plot2, ncol=2)
```


What do you notice in each of the bar charts for mean price by clarity and mean price by color?
clarity: I1 (worst), SI1, SI2, VS1, VS2, VVS1, VVS2, IF (best)
color: J (worst) to D (best)

Generally the better the color, the lower the mean price, which is quite surprising. Similarly, aside from a couple spikes, higher clarity diamonds have lower mean prices.

This is strange. Let's consider the mean price w.r.t. cut
```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))
ggplot(data=diamonds_mp_by_cut, aes(factor(cut), y=mean_price)) + geom_bar(stat="identity")
```


---
Problem Set 5
---
Create a histogram of diamond prices. Facet the histogram by diamond color and 
use cut to color the histogram bars
```{r}
ggplot(data=diamonds, aes(x=price, fill=cut)) + geom_histogram() +
  facet_wrap(~color)
```

Create a scatterplot of diamond price vs.
table and color the points by the cut of
the diamond.
```{r}
ggplot(data=diamonds, aes(x=table, y=price, color=cut)) +
  geom_jitter() + 
  scale_x_continuous(breaks=seq(50, 80, 2), limits=c(50, 80))
```


Create a scatterplot of diamond price vs.
volume (x * y * z) and color the points by
the clarity of diamonds. Use scale on the y-axis
to take the log10 of price. You should also
omit the top 1% of diamond volumes from the plot.
```{r}
diamonds$volume = diamonds$x * diamonds$y * diamonds$z
ggplot(data=subset(diamonds, price < quantile(price, probs=0.9) & volume > 0),
       aes(x=volume, y=price, color=clarity)) + 
  geom_jitter() + 
  scale_y_log10(breaks=c(300, 550, 1000, 1700, 3000, 5500, 10000)) +
  xlim(0, 500)
```


Create a scatter plot of the price/carat ratio
of diamonds. The variable x should be
assigned to cut. The points should be colored
by diamond color, and the plot should be
faceted by clarity.
```{r}
ggplot(data=diamonds, aes(x=cut, y=price/carat)) +
  geom_jitter(aes(color=color), alpha=0.2) + 
  facet_wrap(~ clarity)
```
